---
title: "Scrap power calcs"
output: html_notebook
---

## Summary
This R notebook contains basic formulas for running power calcs, especially for a typical RCT

## Formulas
###  Standard baseline + end line clustered RCT with covariates
This is the most common RCT I encounter.  We randomize at the cluster level, collect data for one baseline and one end line, and run an ANCOVA to estimate ATE.

$$ MDE = M_J\sigma_y\sqrt{\frac{2\rho}{J}+\frac{2(1-\rho)(1-R_K^2)}{JK}} $$
$$ M_J = F^{-1}_t\left(1-\frac{\alpha}{2},J\right) + F^{-1}_t\left(\beta,J\right) $$
Where...

1. $\sigma_y$ is the variance of the outcome variable.
2. $\rho$ is ICC
3. J is the number of clusters in treatment or control (not both). Assumed constant across T and C
4. K is the number of units per cluster. Assumed constant across all clusters.
5. $\alpha$ is significance level (usually .05)
6. $\beta$ is power (usually .8)
7. $R_K^2$ is the R squared from a regression of the outcome on the covariates



```{r}

mj <- function (j, alpha = .05, beta = .8) {
  multiplier <- qt(1-alpha/2, j)+qt(beta,j)
  return(multiplier)
}

mde <- function (j, k, sigmay, icc, rsq, alpha = .05, beta = .8) {
  quant_under_sqrt <- (2*icc/j)+2*(1-icc)*(1-rsq)/(j*k)
  effect <- mj(j, alpha, beta)*sigmay*(quant_under_sqrt^.5)
  return(effect)
}


```

Replicating the APRIGP power calcs
```{r}
villages <- 110
hh_per_v <- 10
icc <- 0.05
rsquared <- 0
sd_cons <- 721

mde(villages, hh_per_v, sd_cons, icc, rsquared)



```

