---
title: "Explore UDISE data"
output: html_notebook
---

### Where I got this data
All of this data is from IDinsight.  Not sure who first got the data, but it seems like Madhav was the one who cleaned it and did most of the analysis.

### Contents of the folder
All data is for academic year 2016-17.  The data is separated by state.  Each state folder contains the following main Excel files:

1. Basic data
    a. Basics of the school including school code, academic year, school name, cluster name, village name
2. DisabledEnrolment
3. FacilityData
    a. Presence of basic facility infrastructure such as water, computers, toilets, etc
4. GeneralData
    a. A little unclear whats in here but it seems like stuff related to school management
5. TotalEnrolment
    a. Unclear, but I think that this contains data on enrollment by gender and class
    b. For example, there is a variable "C1_TOTB" which I think refers to the total class 1 enrollment of boys
    
In addition to these files, there are a number of other excel files (see the output from the list.files command below)

```{r setup}
library(tidyverse)
path <- "C:/Users/dougj/Documents/Data/Education/DISE Data/Dise Data 2016-17 Excel"
```

List the files in the AP state folder
```{r}
list.files(file.path(path, "AP"))
```

